---
title: "Compar"
author: "Kaiiln Lu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(recordlinkR)
library(stringdist)
library(keras)
library(pROC)
library(ggplot2)
```


```{r}
data(iowa_sample)
```


Jaro-Winkler distance is calculated using the stringdist library. Autoencoder distance is the vector of the consine between each encoded pair, normalized to 0-1. Encoded vectors values can be either left as is or binarized. 

```{r}
calculate_jw <- function(vec1, vec2) {
  vec1 <- as.character(vec1)
  vec2 <- as.character(vec2)
  jw <- vector(mode='list', length=length(vec1))
  for (i in 1:length(vec1)) {
    jw[i] <- 1 - stringdist::stringdist(vec1[i], vec2[i], method='jw')
  }
  return (as.numeric(jw))
}

normalize <- function(vec) {
  max.val <- max(vec)
  min.val <- min(vec)
  return ((vec-min.val) / (max.val-min.val))
}

autoencoder_dist <- function(mat1, mat2) {
  ae <- vector(mode='list', length=dim(mat1)[1])
  for (i in 1:dim(mat1)[1]) {
    ae[i] <- mat1[i,] %*% mat2[i,]
  }
  return(normalize(as.numeric(ae)))
}

```

Calculating distance measures on the Iowa last name. 

```{r}
iowa_training_data$jw_lname <-calculate_jw(iowa_training_data$lname1915,
                                           iowa_training_data$lname1940)
```


Load encoder models and encode. 

```{r}
data("encoder_iowa_last_2")
encoder_iowa_last_2 <- unserialize_model(encoder_iowa_last_2)

lname1915_enc_2 <- encodeVector(iowa_training_data$lname1915, encoder_iowa_last_2)
lname1940_enc_2 <- encodeVector(iowa_training_data$lname1940, encoder_iowa_last_2)

data("encoder_iowa_last_4")
encoder_iowa_last_4 <- unserialize_model(encoder_iowa_last_4)

lname1915_enc_4 <- encodeVector(iowa_training_data$lname1915, encoder_iowa_last_4)
lname1940_enc_4 <- encodeVector(iowa_training_data$lname1940, encoder_iowa_last_4)

data("encoder_iowa_last_8")
encoder_iowa_last_8 <- unserialize_model(encoder_iowa_last_8)

lname1915_enc_8 <- encodeVector(iowa_training_data$lname1915, encoder_iowa_last_8)
lname1940_enc_8 <- encodeVector(iowa_training_data$lname1940, encoder_iowa_last_8)

data("encoder_iowa_last_24")
encoder_iowa_last_24 <- unserialize_model(encoder_iowa_last_24)

lname1915_enc_24 <- encodeVector(iowa_training_data$lname1915, encoder_iowa_last_24)
lname1940_enc_24 <- encodeVector(iowa_training_data$lname1940, encoder_iowa_last_24)

data("encoder_iowa_last_48")
encoder_iowa_last_48 <- unserialize_model(encoder_iowa_last_48)

lname1915_enc_48 <- encodeVector(iowa_training_data$lname1915, encoder_iowa_last_48)
lname1940_enc_48 <- encodeVector(iowa_training_data$lname1940, encoder_iowa_last_48)

data("encoder_iowa_last_96")
encoder_iowa_last_96 <- unserialize_model(encoder_iowa_last_96)

lname1915_enc_96 <- encodeVector(iowa_training_data$lname1915, encoder_iowa_last_96)
lname1940_enc_96 <- encodeVector(iowa_training_data$lname1940, encoder_iowa_last_96)

data("encoder_iowa_last_384")
encoder_iowa_last_384 <- unserialize_model(encoder_iowa_last_384)

lname1915_enc_384 <- encodeVector(iowa_training_data$lname1915, encoder_iowa_last_384)
lname1940_enc_384 <- encodeVector(iowa_training_data$lname1940, encoder_iowa_last_384)
```

Calculate autoencoder distances without binarization. 

```{r}
iowa_training_data$lname_enc_2 <- autoencoder_dist(lname1915_enc_2, 
                                                    lname1940_enc_2)

iowa_training_data$lname_enc_4 <- autoencoder_dist(lname1915_enc_4, 
                                                    lname1940_enc_4)

iowa_training_data$lname_enc_8 <- autoencoder_dist(lname1915_enc_8, 
                                                    lname1940_enc_8)

iowa_training_data$lname_enc_24 <- autoencoder_dist(lname1915_enc_24, 
                                                    lname1940_enc_24)

iowa_training_data$lname_enc_48 <- autoencoder_dist(lname1915_enc_48, 
                                                    lname1940_enc_48)

iowa_training_data$lname_enc_96 <- autoencoder_dist(lname1915_enc_48, 
                                                    lname1940_enc_48)

iowa_training_data$lname_enc_384 <- autoencoder_dist(lname1915_enc_48, 
                                                    lname1940_enc_48)
```


Calculate ROC and AUC for JW and 
```{r}
roc_lname_jw <- roc(iowa_training_data$match, iowa_training_data$jw_lname)
paste('jw AUC =', round(auc(roc_lname_jw),4)) 

roc_lname_enc_2 <- roc(iowa_training_data$match, iowa_training_data$lname_enc_2)
paste('LSTM 2 AUC =', round(auc(roc_lname_enc_2),4)) 

roc_lname_enc_4 <- roc(iowa_training_data$match, iowa_training_data$lname_enc_4) 
paste('LSTM 4 AUC =', round(auc(roc_lname_enc_4),4))

roc_lname_enc_8 <- roc(iowa_training_data$match, iowa_training_data$lname_enc_8) 
paste('LSTM 8 AUC =', round(auc(roc_lname_enc_8),4)) 

roc_lname_enc_24 <- roc(iowa_training_data$match, iowa_training_data$lname_enc_24) 
paste('LSTM 24 AUC =', round(auc(roc_lname_enc_24),4)) 

roc_lname_enc_48 <- roc(iowa_training_data$match, iowa_training_data$lname_enc_48) 
paste('LSTM 48 AUC =', round(auc(roc_lname_enc_48),4)) 

roc_lname_enc_96 <- roc(iowa_training_data$match, iowa_training_data$lname_enc_96) 
paste('LSTM 96 AUC =', round(auc(roc_lname_enc_96),4)) 

roc_lname_enc_384 <- roc(iowa_training_data$match, iowa_training_data$lname_enc_384) 
paste('LSTM 384 AUC =', round(auc(roc_lname_enc_384),4))
```


```{r}
g <- ggroc(list('jw AUC=.8961'=roc_lname_jw, 
                'enc002 AUC=.7413'=roc_lname_enc_2, 
                'enc004 AUC=.7211'=roc_lname_enc_4, 
                'enc008 AUC=.6778'=roc_lname_enc_8, 
                'enc024 AUC=.7504'=roc_lname_enc_24, 
                'enc048 AUC=.8363'=roc_lname_enc_48, 
                'enc096 AUC=.8363'=roc_lname_enc_96, 
                'enc384 AUC=.8363'=roc_lname_enc_384))

g + ggtitle('Iowa Last Name - Vector Distance')
```


Calculate autoencoder distance binarized. 

To binarize, compare each encoded vector to the median of all encoded vectors from one columns. In this case, the median is calculated using the 1915 last name. 

```{r}
lname1915_enc_2_list <- binarizeVector(lname1915_enc_2)
lname1915_enc_2_b <- lname1915_enc_2_list[[1]]
lname1940_enc_2_b <- binarizeVector(lname1940_enc_2, lname1915_enc_2_list[[2]])[[1]]

iowa_training_data$lname_enc_2_b <- autoencoder_dist(lname1915_enc_2_b, 
                                                    lname1940_enc_2_b)

lname1915_enc_4_list <- binarizeVector(lname1915_enc_4)
lname1915_enc_4_b <- lname1915_enc_4_list[[1]]
lname1940_enc_4_b <- binarizeVector(lname1940_enc_4, lname1915_enc_4_list[[2]])[[1]]

iowa_training_data$lname_enc_4_b <- autoencoder_dist(lname1915_enc_4_b, 
                                                    lname1940_enc_4_b)

lname1915_enc_8_list <- binarizeVector(lname1915_enc_8)
lname1915_enc_8_b <- lname1915_enc_8_list[[1]]
lname1940_enc_8_b <- binarizeVector(lname1940_enc_8, lname1915_enc_8_list[[2]])[[1]]

iowa_training_data$lname_enc_8_b <- autoencoder_dist(lname1915_enc_8_b, 
                                                    lname1940_enc_8_b)

lname1915_enc_8_list <- binarizeVector(lname1915_enc_8)
lname1915_enc_8_b <- lname1915_enc_8_list[[1]]
lname1940_enc_8_b <- binarizeVector(lname1940_enc_8, lname1915_enc_8_list[[2]])[[1]]

iowa_training_data$lname_enc_8_b <- autoencoder_dist(lname1915_enc_8_b, 
                                                    lname1940_enc_8_b)

lname1915_enc_24_list <- binarizeVector(lname1915_enc_24)
lname1915_enc_24_b <- lname1915_enc_24_list[[1]]
lname1940_enc_24_b <- binarizeVector(lname1940_enc_24, lname1915_enc_24_list[[2]])[[1]]

iowa_training_data$lname_enc_24_b <- autoencoder_dist(lname1915_enc_24_b, 
                                                    lname1940_enc_24_b)

lname1915_enc_48_list <- binarizeVector(lname1915_enc_48)
lname1915_enc_48_b <- lname1915_enc_48_list[[1]]
lname1940_enc_48_b <- binarizeVector(lname1940_enc_48, lname1915_enc_48_list[[2]])[[1]]

iowa_training_data$lname_enc_48_b <- autoencoder_dist(lname1915_enc_48_b, 
                                                    lname1940_enc_48_b)


lname1915_enc_96_list <- binarizeVector(lname1915_enc_96)
lname1915_enc_96_b <- lname1915_enc_96_list[[1]]
lname1940_enc_96_b <- binarizeVector(lname1940_enc_96, lname1915_enc_96_list[[2]])[[1]]

iowa_training_data$lname_enc_96_b <- autoencoder_dist(lname1915_enc_96_b, 
                                                    lname1940_enc_96_b)

lname1915_enc_384_list <- binarizeVector(lname1915_enc_384)
lname1915_enc_384_b <- lname1915_enc_384_list[[1]]
lname1940_enc_384_b <- binarizeVector(lname1940_enc_384, lname1915_enc_384_list[[2]])[[1]]

iowa_training_data$lname_enc_384_b <- autoencoder_dist(lname1915_enc_384_b, 
                                                    lname1940_enc_384_b)
```

```{r}
roc_lname_enc_2 <- roc(iowa_training_data$match,
                       iowa_training_data$lname_enc_2_b)
paste('LSTM 2 AUC =', round(auc(roc_lname_enc_2),4)) 

roc_lname_enc_4 <- roc(iowa_training_data$match, 
                       iowa_training_data$lname_enc_4_b) 
paste('LSTM 4 AUC =', round(auc(roc_lname_enc_4),4))

roc_lname_enc_8 <- roc(iowa_training_data$match,
                       iowa_training_data$lname_enc_8_b) 
paste('LSTM 8 AUC =', round(auc(roc_lname_enc_8),4)) 

roc_lname_enc_24 <- roc(iowa_training_data$match,
                        iowa_training_data$lname_enc_24_b) 
paste('LSTM 24 AUC =', round(auc(roc_lname_enc_24),4)) 

roc_lname_enc_48 <- roc(iowa_training_data$match,
                        iowa_training_data$lname_enc_48_b) 
paste('LSTM 48 AUC =', round(auc(roc_lname_enc_48),4)) 

roc_lname_enc_96 <- roc(iowa_training_data$match, 
                        iowa_training_data$lname_enc_96_b) 
paste('LSTM 96 AUC =', round(auc(roc_lname_enc_96),4)) 

roc_lname_enc_384 <- roc(iowa_training_data$match,
                         iowa_training_data$lname_enc_384_b) 
paste('LSTM 384 AUC =', round(auc(roc_lname_enc_384),4))
```


```{r}
g <- ggroc(list('jw AUC=.8961'=roc_lname_jw, 
                'enc002 AUC=.6343'=roc_lname_enc_2, 
                'enc004 AUC=.6741'=roc_lname_enc_4, 
                'enc008 AUC=.7193'=roc_lname_enc_8, 
                'enc024 AUC=.8111'=roc_lname_enc_24, 
                'enc048 AUC=.8699'=roc_lname_enc_48, 
                'enc096 AUC=.8813'=roc_lname_enc_96, 
                'enc384 AUC=.9078'=roc_lname_enc_384))

g + ggtitle('Iowa Last Name - Binarized Vector Distance')

```

Jaro Winkler and Binary LSTM vector distance can be combined by taking a product. 

```{r}
iowa_training_data$lname_jw_enc_2 <- iowa_training_data$lname_enc_2 * iowa_training_data$jw_lname

iowa_training_data$lname_jw_enc_4 <- iowa_training_data$lname_enc_4 * iowa_training_data$jw_lname

iowa_training_data$lname_jw_enc_8 <- iowa_training_data$lname_enc_8 * iowa_training_data$jw_lname

iowa_training_data$lname_jw_enc_24 <- iowa_training_data$lname_enc_24 * iowa_training_data$jw_lname

iowa_training_data$lname_jw_enc_48 <- iowa_training_data$lname_enc_48 * iowa_training_data$jw_lname

iowa_training_data$lname_jw_enc_96 <- iowa_training_data$lname_enc_96 * iowa_training_data$jw_lname

iowa_training_data$lname_jw_enc_384 <- iowa_training_data$lname_enc_384 * iowa_training_data$jw_lname
```


```{r}
roc_lname_enc_2 <- roc(iowa_training_data$match,
                       iowa_training_data$lname_jw_enc_2)
paste('LSTM 2 AUC =', round(auc(roc_lname_enc_2),4)) 

roc_lname_enc_4 <- roc(iowa_training_data$match, 
                       iowa_training_data$lname_jw_enc_4) 
paste('LSTM 4 AUC =', round(auc(roc_lname_enc_4),4))

roc_lname_enc_8 <- roc(iowa_training_data$match,
                       iowa_training_data$lname_jw_enc_8) 
paste('LSTM 8 AUC =', round(auc(roc_lname_enc_8),4)) 

roc_lname_enc_24 <- roc(iowa_training_data$match,
                        iowa_training_data$lname_jw_enc_24) 
paste('LSTM 24 AUC =', round(auc(roc_lname_enc_24),4)) 

roc_lname_enc_48 <- roc(iowa_training_data$match,
                        iowa_training_data$lname_jw_enc_48) 
paste('LSTM 48 AUC =', round(auc(roc_lname_enc_48),4)) 

roc_lname_enc_96 <- roc(iowa_training_data$match, 
                        iowa_training_data$lname_jw_enc_96) 
paste('LSTM 96 AUC =', round(auc(roc_lname_enc_96),4)) 

roc_lname_enc_384 <- roc(iowa_training_data$match,
                         iowa_training_data$lname_jw_enc_384) 
paste('LSTM 384 AUC =', round(auc(roc_lname_enc_384),4))
```

# Blocking 

```{r}
cols.encoder <- list('A'='lname1915', 'B'='lname1940')

blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_2, 
                encoder.block.method = 'cluster', 
                encoder.nclusters = 10,
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```

```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_4, 
                encoder.block.method = 'cluster', 
                encoder.nclusters = 10,
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```


```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_8, 
                encoder.block.method = 'cluster', 
                encoder.nclusters = 10,
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```


```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_24, 
                encoder.block.method = 'cluster', 
                encoder.nclusters = 10,
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```


```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_48, 
                encoder.block.method = 'cluster', 
                encoder.nclusters = 10,
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```


```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_96, 
                encoder.block.method = 'cluster', 
                encoder.nclusters = 10,
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```

```{r}
cluster_blocks <- data.frame(dim.latent=c("2", "4", "8", "24", "48", "96"), 
                             proportion.comparisons=c(14.34, 15.51, 11.37, 11.36, 10.36 ,11.14), 
                             pairs.completeness=c(85.02, 85.39, 86.71, 90.90, 90.51, 90.16))
cluster_blocks$dim.latent <- factor(cluster_blocks$dim.latent, levels=cluster_blocks$dim.latent)

g <- ggplot(data=cluster_blocks, aes(x=dim.latent, y=proportion.comparisons)) + 
  geom_bar(stat='identity') +
  geom_text(aes(label=proportion.comparisons), vjust=1.6, color="white", size=3.5)+
  ggtitle('Proportion of Comparisons After Blocking on Encoded Clusters (n.clusters=10)') + 
  theme_minimal() 
g
```

```{r}
g <- ggplot(data=cluster_blocks, aes(x=dim.latent, y=pairs.completeness)) + 
  geom_bar(stat='identity') +
  geom_text(aes(label=pairs.completeness), vjust=1.6, color="white", size=3.5)+
  ggtitle('Pairs Completeness After Blocking on Encoded Clusters (n.clusters=10)') + 
  theme_minimal() 
g
```


Blocking using 96 dim model, different number of clusters. 

```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_96, 
                encoder.block.method = 'cluster', 
                encoder.nclusters = 2,
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```


```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_96, 
                encoder.block.method = 'cluster', 
                encoder.nclusters = 4,
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```

```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_96, 
                encoder.block.method = 'cluster', 
                encoder.nclusters = 6,
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```

```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_96, 
                encoder.block.method = 'cluster', 
                encoder.nclusters = 8,
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```

```{r}
cluster_blocks_96 <- data.frame(n.clusters=c("2", "4", "6", "8", "10"), 
                             proportion.comparisons=c(59.76, 28.33, 17.28, 13.44, 11.14), 
                             pairs.completeness=c(96.2, 93.08, 91.04, 90.65, 90.16))

cluster_blocks_96$n.clusters <- factor(cluster_blocks_96$n.clusters, levels=cluster_blocks_96$n.clusters)

g <- ggplot(data=cluster_blocks_96, aes(x=n.clusters, y=proportion.comparisons)) + 
  geom_bar(stat='identity') +
  geom_text(aes(label=proportion.comparisons), vjust=1.6, color="white", size=3.5)+
  ggtitle('Proportion of Comparisons After Blocking on \n Clustered Encoded Vecs (dim.latent=96)') + 
  theme_minimal() 
g
```

```{r}
g <- ggplot(data=cluster_blocks_96, aes(x=n.clusters, y=pairs.completeness)) + 
  geom_bar(stat='identity') +
  geom_text(aes(label=pairs.completeness), vjust=1.6, color="white", size=3.5)+
  ggtitle('Pairs Completeness After Blocking on \n Clustered Encoded Vecs (dim.latent=96)') + 
  theme_minimal() 
g
```


Blocking on binary. 

```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_2, 
                encoder.block.method = 'binary', 
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```


```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_4, 
                encoder.block.method = 'binary', 
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```




```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_8, 
                encoder.block.method = 'binary', 
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```



```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_24, 
                encoder.block.method = 'binary', 
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```


```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_48, 
                encoder.block.method = 'binary', 
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```


```{r}
blocks <- block(iowa_1915, iowa_1940, 
                cols.encoder=cols.encoder, 
                encoder.model.path=encoder_iowa_last_96, 
                encoder.block.method = 'binary', 
                n.cores = 1, 
                known.matches = iowa_true_matches[,1:2])
```



```{r}
binary_blocks <- data.frame(dim.latent=c("2", "4", "8", "24", "48", "96"), 
                             proportion.comparisons=c(49.48, 35.49, 6.06, 0.127, 0.121, 0.129), 
                             paris.completeness=c(93.89, 91.02, 81.9, 73.31,72.15,71.83))

binary_blocks$dim.latent <- factor(binary_blocks$dim.latent, levels=binary_blocks$dim.latent)

g <- ggplot(data=binary_blocks, aes(x=dim.latent, y=proportion.comparisons)) + 
  geom_bar(stat='identity') +
  geom_text(aes(label=proportion.comparisons), vjust=-.3, color="gray", size=3.5)+
  ggtitle('Proportion of Comparisons After Blocking on Binarized Encoded Vectors') + 
  theme_minimal() 
g
```


```{r}
binary_blocks <- data.frame(dim.latent=c("2", "4", "8", "24", "48", "96"), 
                             proportion.comparisons=c(49.48, 35.49, 6.06, 0.127, 0.121, 0.129), 
                             pairs.completeness=c(93.89, 91.02, 81.9, 73.31,72.15,71.83))

binary_blocks$dim.latent <- factor(binary_blocks$dim.latent, levels=binary_blocks$dim.latent)

g <- ggplot(data=binary_blocks, aes(x=dim.latent, y=pairs.completeness)) + 
  geom_bar(stat='identity') +
  geom_text(aes(label=pairs.completeness), vjust=-.3, color="gray", size=3.5)+
  ggtitle('Proportion of Comparisons After Blocking on Binarized Encoded Vectors') + 
  theme_minimal() 
g
```



