---
title: "Preprocessing Data"
author: "Kailin Lu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(recordlinkR)
```


## An Example Dataset 

Datasets with name fields can come in a variety of formats. some datasets may provide full names whereas others may have separate columns for first and last. Full names can be listed in order of <FIRST, LAST> or <LAST, FIRST>. Additionally, there may be inconsistencies in how names are capitalized and whether punctuation is included. 

Below is an example of a dataset with several name columns following different formats. 

```{r}
# Some sample character and numeric data 
names <- c('Doe, Jon', 'GRAY, Amy**', 'JACKSON, MICHAEL!', 'smith, jane')
mothers.names <- c('Dorian\' Do', 'Brittany John-son Smith', 'Mary JACKSON', 'Sam')
nicknames <- c('JON', 'AMY', 'MIKE', 'JAN')
age <- c(15, 14, 12, 17)

# Create dataframe
data <- data.frame(names, mothers.names, nicknames, age, stringsAsFactors = FALSE)

knitr::kable(data, caption = 'Sample Dataframe')
```

The function to preprocess a dataframe is `cleanNames`. 

To apply changes to the variables names, mothers.names, and nicknames, add as a vector to the `cols` parameter of cleanNames. By default, `lower` and `strip.punctuation` are TRUE. 

```{r}
data.cleaned <- cleanNames(data,
                           cols = c('names', 'mothers.names', 'nicknames'),
                           lower = T,
                           strip.punctuation = T)
knitr::kable(data.cleaned, caption = 'Cleaned Sample Dataframe')
```

To train an encoder model on names it is helpful to train models on first or last names alone, rather than in conjunction. 
String columns can be split using `cleanNames` by specifying a character split on and names of the newly created variables. 


```{r}
split.pattern <- list('names' = ',', 'mothers.names' = ' ')
new.col.names <- list('names' = c('last', 'first'), 'mothers.names' = c('mothers.first', 'mothers.last'))
  
data.cleaned.and.split <- cleanNames(data,
                                     cols = c('names', 'mothers.names', 'nicknames'),
                                     split.pattern = split.pattern, 
                                     split.col.names = new.col.names, 
                                     lower = T,
                                     strip.punctuation = T)

knitr::kable(data.cleaned.and.split, caption = 'Cleaned Sample Dataframe with Split Columns')
```







