---
title: "Preprocessing Data"
author: "Kailin Lu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(recordlinkR)
```


## An Example Dataset 

Datasets with name fields can come in a variety of formats. some datasets may provide full names whereas others may have separate columns for first and last. Full names can be listed in order of <FIRST, LAST> or <LAST, FIRST>. Additionally, there may be inconsistencies in how names are capitalized and whether punctuation is included. 

Below is an example of a dataset with several name columns following different formats. 

```{r}
# Some sample character and numeric data 
names <- c('Doe, Jon', 'GRAY, Amy**', 'JACKSON, MICHAEL', 'smith, jane')
mothers.names <- c('Dorian\' Do', 'Brittany John-son Smith', 'Mary JACKSON', 'Sam')
nicknames <- c('JON', 'AMY', 'MIKE', 'JAN')
age <- c(15, 14, 12, 17)

# Create dataframe
data <- data.frame(names, mothers.names, nicknames, age, stringsAsFactors = FALSE)

knitr::kable(data, caption = 'Sample Dataframe')
```

```{r}
data.cleaned <- cleanNames(data,
                           cols = c('names', 'mothers.names', 'nicknames'),
                           split = T,
                           split.pattern = c(',', ' '),
                           new.col.names = list('names' = c('last', 'first'),
                                                'mothers.names' = c('first.m', 'last.m')),
                           lower = T,
                           strip.punctuation = T)

knitr::kable(data.cleaned, caption = 'Cleaned Sample Dataframe')
```
