---
title: "Training Encoding Models"
author: "Kailin Lu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(recordlinkR)

# Load sample Iowa data
data(iowa_sample)
```


This chunk is set to currently not run for convenience. Default parameters of the model are likely 
not sufficient to achieve a good model fit. 

To see parameters and check default settings on the model, see the help manual at `?trainModel`. 

```{r eval=FALSE}
namesA <- iowa_1915$lname1915[iowa_true_matches$index1915]
namesB <- iowa_1940$lname1940[iowa_true_matches$index1940]

encoder <- trainModel(namesA = namesA, 
                      namesB = namesB, 
                      dim.encode = 64, 
                      dim.decode = 64,
                      dim.latent = 48, 
                      num.encode.layers = 2, 
                      num.decode.layers = 2, 
                      earlystop.patience = 20, 
                      batch.size = 32,
                      epochs = 350, 
                      lr = 5e-4, 
                      save.dir="~/blocking_models/iowa_last_48/")
```

```{r}
encoder
```



```{r}
save(encoder_iowa_last_48, file="encoder_iowa_last_48.rda")
```

Outside of model accuracy, models can be evaluated by plotting a sample of names represented as a 
2-dimensional projection of the latent space. 

```{r}
# Load prepackaged sample encoders 
loadSampleEncoders()

# Plot sample of 100 first names 
plotNames(names=iowa_1915$fname1915, 
          n.samples = 100, 
          encoder.model.path = encoder_iowa_first_512) 
```



























